---
title: "SRT411 Assignment 0"
author: "Christopher Gow"
date: "February 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#SRT411 Assignment 0
### Introduction
My name is Christopher Gow, student number 102780152.  My username on GitHub is cgow1.  The task for this assignment is to complete the To Do sections on a document called A (Very) Short Introduction to R by Paul Torfs & Claudia Brauer.  The link to the document can be found at this web address: https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf .  

The purpose of this assignment is to give a brief introduction to R.  Throughout this document there are To Do sections which test you as you are progressing throughout the document.  After the document explains a section of R, it provides you with a To Do to make sure you understand what was being taught to you.  


###ToDo 1 
Compute the difference between 2014 and the year you started at this university and divide this by the difference between 2014 and the year you were born. Multiply this with 100 to get the percentage of your life you have spent at this university. Use brackets if you need them.
```{r}
((2017-2015)/(2017-1997)) * 100
```

###ToDo 2 
Repeat the previous ToDo, but with several steps in between. You can give the variables
any name you want, but the name has to start with a letter.

```{r}
life = (2017-2015)
dob = (2017-1997)
life / dob * 100
```

###ToDo 3 
Compute the sum of 4, 5, 8 and 11 by first combining them into a vector and then using the
function sum.
```{r}
x = c(4,5,8,11)
sum (x)
```

###ToDo 4 
Plot 100 normal random numbers.
```{r}
plot(rnorm(100))
```

###ToDo 5 
Find help for the sqrt function
```{r}
help(sqrt)
```

###ToDo 6 
Make a file called firstscript.R containing Rcode that generates 100 random numbers and
plots them, and run this script several times.
```{r}
source('C:/Users/Chris/Desktop/firstscript.R')
```

###ToDo 7 
Put the numbers 31 to 60 in a vector named P and in a matrix with 6 rows and 5 columns
named Q. Tip: use the function seq. Look at the different ways scalars, vectors and matrices
are denoted in the workspace window
```{r}
P = seq (from = 31, to = 60, by = 1)
Q = matrix(P, nrow = 6, ncol = 5)
P
Q
```

###ToDo 8 
Make a script file which constructs three random normal vectors of length 100. Call these
vectors x1, x2 and x3. Make a data frame called t with three columns (called a, b and c) containing respectively x1, x1+x2 and x1+x2+x3. Call the following functions for this data frame:
plot(t) and sd(t). Can you understand the results? Rerun this script a few times.
```{r}
x1 = seq (from = 1, to = 100, by = 1)
x2 = seq (from = 101, to = 200, by = 1)
x3 = seq (from = 201, to = 300, by = 1)
t = data.frame(a = x1, b = x1 + x2, c = x1 + x2 + x3)
plot(t)
help(rgb)
```

###ToDo 9 
Add these lines to the script file of the previous section. Try to find out, either by experimenting or by using the help, what the meaning is of rgb, the last argument of rgb, lwd, pch, cex.
```{r}
plot(t$a, type="l", ylim=range(t), lwd=3, col=rgb(1,0,0,0.3))
lines(t$b, type="s", lwd=2, col=rgb(0.3,0.4,0.3,0.9))
points(t$c, pch=20, cex=4, col=rgb(0,0,1,0.3))
```

###ToDo 10 
Make a file called tst1.txt in Notepad from the example in Figure 4 and store it in your
working directory. Write a script to read it, to multiply the column called g by 5 and to store
it as tst2.txt.
```{r}
d = data.frame(a = c(3,4,5), b = c(12,43,54))
write.table(d, file="tst1.txt", row.names=FALSE)
write.table(d, file = "tst2.txt", row.names=FALSE)
d2 = read.table(file = "tst2.txt", header = TRUE)
d2$a*5
```

###ToDo 11
Compute the mean of the square root of a vectorof 100 random numbers. What happens?
```{r}
sqrt(mean(rnorm(100)))
```
Somtimes you will receive an error that says NaNs produced.

###ToDo 12 
Make a graph with on the x-axis: today, Sinterklaas 2014 and your next birthday and on
the y-axis the number of presents you expect on each of these days. Tip: make two vectors first.
```{r}
date = strptime( c("20170130", "20170512", "20171205"), format = "%Y%m%d")
presents = c(1 , 0, 10)
plot(date,presents, xlab = "Dates", ylab = "Presents expected")
```

###ToDo 13 
Make a vector from 1 to 100. Make a for-loop which runs through the whole vector. Multiply
the elements which are smaller than 5 and larger than 90 with 10 and the other elements with 0.1.
```{r}
forloop = seq(from = 1, to = 100, by = 1)
s = c()
for(i in 1:100)
{
if (forloop[i]<5)
  {
  s[i] = forloop[i] * 5;
  }
   else if (forloop[i] > 90)
  {
   s[i] = forloop[i] * 10;
   }
  else
   {
   s[i] = forloop[i] * 0.1;
   }
   }
s
```

###ToDo 14
Write a function for the previous ToDo, so that you can feed it any vector you like
(as argument). Use a for-loop in the function to do the computation with each element.
Use the standard R function length in the specification of the counter.
```{r}
userinput = function(arg1)
   {
     len = length(arg1)
     for(i in 1:len)
       {
         if (arg1[i] < 5)
           {
             arg1[i] = arg1[i] * 5;
             }
         else if (arg1[i] > 90)
           {
             arg1[i] = arg1[i] * 10;
             }
         else
           {
             arg1[i] = arg1[i] * 0.1;
             }
         }
     return (arg1)
}
userinput(arg1 = 20)
```

###ToDo 15
The ToDo above can be done more easily
and quickly without a for-loop but with regular vectorcomputations.
```{r}
userinput2 = function(arg1)
   {
     if (arg1 < 5)
       {
         arg1 = arg1 * 5;
         }
     else if (arg1 > 90)
       {
         arg1 = arg1 * 10;
         }
     else
       {
         arg1 = arg1 * 0.1;
         }
     return (arg1)
}
userinput2(arg1 = 10)
```

##Sources Used
A (very) short introduction to R
https://cran.r-project.org/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf

R Markdown - Dynamic Documents for R
http://rmarkdown.rstudio.com/

Writing reproducible reports in R with markdown, knitr and pandoc
https://nicercode.github.io/guides/reports/

Markdown
http://kbroman.org/knitr_knutshell/pages/markdown.html

knitr with R Markdown
http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html

R markdown cheatsheet
https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf